---
title: "Task 2 — Response Density"
subtitle: "Verasight Data Scientist Case Study"
format:
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
    theme: litera
    highlight-style: github
    code-fold: true
    code-summary: "Show code"
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false
library(here)
source(here::here("utils", "read_survey.R"))
source(here::here("utils", "density.R"))
library(dplyr)
library(ggplot2)
library(plotly)

# Load + clean (de-duplicate exact rows, drop unattributable NA-ID rows).
raw <- read_survey(here::here("data", "full-response-db.rds"), quiet = TRUE)
n_na <- sum(is.na(raw$ID)); n_dupe <- sum(duplicated(raw))
db   <- clean_responses(raw)
per_user <- db %>% count(ID, name = "responses")

# Task 2.1
share_50 <- min_user_share_for(per_user$responses, 0.50)
n_users  <- nrow(per_user)

# All-registered contrast
users_all <- read_survey(here::here("data", "users.rds"), quiet = TRUE)
n_zero    <- n_distinct(users_all$ID) - n_users
share_50_allreg <- min_user_share_for(c(per_user$responses, rep(0, n_zero)), 0.50)

# Task 2.2 — registered in the last 90 days (reference = latest date in the data)
ref_date <- as.Date(max(raw$response_date, na.rm = TRUE))
cutoff   <- ref_date - 90
users_all$signup <- as.Date(users_all$signup_date, format = "%m/%d/%Y")
recent_ids <- users_all$ID[!is.na(users_all$signup) & users_all$signup >= cutoff]
per_recent <- db %>% filter(ID %in% recent_ids) %>% count(ID, name = "responses")
share_50_recent <- min_user_share_for(per_recent$responses, 0.50)

pct <- function(x) sprintf("%.1f%%", 100 * x)
```

> **Task 2 — Response density.** Measure how concentrated survey responses are among
> panelists. (1) Out of all users in the database, what is the smallest percent of users who
> account for 50% of all responses? (2) The same metric, restricted to users who registered
> in the last 90 days. (3) Visualize the full concentration curve — cumulative % of responses
> vs. cumulative % of users — with the 10%–90% thresholds marked.

## Approach

"Density" here is a **concentration** question, so I treat it as a **Lorenz curve**: sort
users from most to least active, accumulate their responses, and read off what share of users
it takes to reach a given share of responses. The metric for parts 1–2 is the *smallest* share
of users (most active first) whose responses reach 50% of the total. All logic lives in a
small reusable helper, [`utils/density.R`](../utils/density.R).

## Data preparation

Before counting, two clean-ups materially change the answer:

- **Missing IDs.** `r sprintf("%.0f%%", 100*n_na/nrow(raw))` of rows (`r format(n_na, big.mark=",")`)
  have **no user ID**. These are unattributable to any panelist, so they are **excluded** —
  a naive `count(ID)` would otherwise lump them into one phantom "user" holding ~22% of all
  responses.
- **Duplicates.** `r format(n_dupe, big.mark=",")` exact-duplicate rows (same ID / survey /
  date) are de-duplicated.

**Which users count as the denominator?** "Users in the database" is ambiguous, and the choice
changes the answer:

- **Responders** — the `r format(n_users, big.mark=",")` users who submitted at least one
  response. **This is the universe I use.**
- **All registered users** — all `r format(n_distinct(users_all$ID), big.mark=",")` accounts,
  including `r format(n_zero, big.mark=",")` who have *never* responded.

Adding those tens of thousands of zero-response accounts makes responses look far more
concentrated — the same 50% is then reached by only **`r pct(share_50_allreg)`** of accounts,
versus **`r pct(share_50)`** among responders. I report the responder-based figure and state
the choice explicitly so the number isn't read out of context.

## Results

| Question | Answer |
|---|---|
| **2.1** Smallest % of users for 50% of responses | **`r pct(share_50)`** (`r format(ceiling(share_50*n_users), big.mark=",")` of `r format(n_users, big.mark=",")`), Gini `r sprintf("%.2f", gini(per_user$responses))` |
| **2.2** Same, registered in the last 90 days | **`r pct(share_50_recent)`** (`r ceiling(share_50_recent*nrow(per_recent))` of `r format(nrow(per_recent), big.mark=",")`), Gini `r sprintf("%.2f", gini(per_recent$responses))` |

*Task 2.2 reference date = the latest date in the data (`r format(ref_date)`); therefore, the
90-day cohort comprises registrations on or after `r format(cutoff)` through
`r format(ref_date)`.*

**Reading the result:** responses are moderately concentrated — the most active
`r pct(share_50)` of responders produce half of all responses. **New panelists are much less
concentrated** (`r pct(share_50_recent)` needed for 50%): concentration builds with tenure as
heavy responders accumulate history.

## Concentration curves

Each curve plots cumulative % of responses (y) against cumulative % of users, most-active-first
(x). The dashed diagonal is perfect equality; the further the navy curve bows above it, the more
concentrated. Teal dots mark the 10%–90% thresholds; the **red dot + guide lines mark the 50%
answer**. Hover for exact values.

### All responders

```{r}
#| label: fig-density-all
#| fig-cap: "Response concentration across all responders (most active first). The dashed diagonal is perfect equality; the red point marks the 50% threshold (the base question)."
ggplotly(density_figure(per_user$responses, "Response density — all responders"),
         tooltip = "text")
```

### Registered in the last 90 days

```{r}
#| label: fig-density-recent
#| fig-cap: "Response concentration among users registered in the last 90 days — visibly less concentrated than the full panel (the curve sits closer to the equality diagonal)."
ggplotly(density_figure(per_recent$responses, "Response density — registered in the last 90 days"),
         tooltip = "text")
```

## Assumptions & edge cases

- **File:** brief said `response-db.rds`; actual file is `data/full-response-db.rds`.
- **Unattributable rows** (missing ID) and **exact duplicates** are dropped (above).
- **Universe** = responders; the all-registered alternative is reported for contrast.
- **"Last 90 days"** is relative to the latest date in the data, not today.
- **Ties / zero-response users:** users are ordered most-active-first and the cumulative curve
  is monotonic, so the smallest user count reaching a target is well defined; zero-response
  users appear only in the all-registered contrast.
