---
title: "Task 1 — Weighted Crosstab Function"
subtitle: "Verasight Data Scientist Case Study"
format:
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
    theme: litera
    highlight-style: github
    code-fold: true
    code-summary: "Show code"
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false
library(here)
source(here::here("utils", "read_survey.R"))          # read_survey()
source(here::here("utils", "weighted_crosstab.R"))    # weighted_crosstab()
library(survey)
library(gt)
library(tibble)
```

> **Task 1 — Construct and run a tabulation function.** Create a single, documented, reusable
> R function (plus any small helpers it needs) to produce weighted proportion tables for each
> survey outcome by each demographic variable in the recent response dataset
> (`2024-054_responses.rds`), with the question wording attached. These data are from a
> nationally representative survey of U.S. adults. Your function should be constructed such
> that a teammate could understand and use your function on a different Verasight survey
> without editing its internals.

## What I built

A single reusable function, `weighted_crosstab()` (defined in
[`utils/weighted_crosstab.R`](../utils/weighted_crosstab.R)), that produces **weighted
proportion tables** — one for each survey **outcome × demographic** — with each question's
wording attached above its table. The execution script
[`code/01_task1_crosstabs.R`](../code/01_task1_crosstabs.R) runs it on survey `2024-054`.

The design goal was the grading bar: *a teammate can run it on a different Verasight survey
without editing any internals.* Everything survey-specific is an **argument**, not hardcoded.

## How the function works

- **Weighting.** These are complex-survey weights (verified: all non-integer, most < 1,
  sum ≈ *n*), not frequency weights. Estimation uses the **`survey`** package —
  `svydesign(ids = ~1, weights = ~weight)` → `svytable()` — then normalizes weighted counts
  **within each demographic column** so each column sums to 100% across the response levels.
- **Question wording** is joined from the reference file (`variable → label`) and shown as
  each table's title.
- **Reusability by inference + override.** By default it *infers* roles — outcomes are columns
  matching `^q[0-9]+$`; demographics are the remaining **categorical** columns (excluding the
  id and weight). Every role is also an explicit argument (`outcomes`, `demographics`,
  `weight_col`, `id_col`, `reference`, `ref_var`/`ref_label`, `outcome_pattern`), and explicit
  arguments always win.
- **Robustness.** Missing demographics are dropped (or shown as `(Missing)` via
  `na_demographic = "keep"`); missing responses are excluded so percentages are among
  answerers; empty response levels are retained at 0%; **continuous / high-cardinality
  columns are hard-skipped** (`max_categories`, default 20) so a variable like raw `age`
  can never explode into dozens of columns; and **negative weights error** (survey weights
  must be ≥ 0).

## Example: survey 2024-054

```{r real-example}
responses <- read_survey(here::here("data", "2024-054_responses.rds"), quiet = TRUE)
reference <- read_survey(here::here("data", "2024-054_reference.rds"), quiet = TRUE)

tables <- weighted_crosstab(responses, reference = reference)

# 8 outcomes (q28–q35) × 6 demographics = 48 tables. Show the accuracy-by-age table:
tables[["q28__age_group4"]]$gt
```

## Manual validation

To confirm the numbers are right — and that the **`survey` package** (not an unweighted
count) is doing the work — I independently recompute **one cell** of the table above by two
other paths and compare:

```{r manual-validation}
#| code-fold: false
tab     <- tables[["q28__age_group4"]]$data
cell_fn <- tab[tab$Response == "Extremely accurate", "18-29", drop = TRUE]     # function output

# Path A — base R weighted.mean() (no survey package at all)
sub        <- responses[!is.na(responses$age_group4) &
                          responses$age_group4 == "18-29" & !is.na(responses$q28), ]
cell_wmean <- weighted.mean(sub$q28 == "Extremely accurate", sub$weight)

# Path B — survey::svymean() on an independent design (a different survey-package call)
d        <- svydesign(ids = ~1, weights = ~weight, data = sub)
cell_svy <- as.numeric(svymean(~ I(q28 == "Extremely accurate"), d)[2])

data.frame(
  method   = c("weighted_crosstab()", "base R weighted.mean()", "survey::svymean()"),
  estimate = round(c(cell_fn, cell_wmean, cell_svy), 7)
)
```

```{r manual-validation-check}
#| code-fold: false
stopifnot(max(abs(c(cell_fn - cell_wmean, cell_fn - cell_svy))) < 1e-9)   # hard check
"All three methods agree to within 1e-9."
```

Inside `weighted_crosstab()` the estimation is done by `survey::svydesign()` +
`survey::svytable()`, so this agreement confirms the weights are applied through the survey
design — not a plain frequency count.

## It works on a different colleague's survey

To show portability, here is a **synthetic survey a colleague might bring** — completely
different column names (`user_id`, `wt`, `item_1`/`item_2`, `region2`, `agegrp`) and a
differently-named reference. The function is driven **entirely by arguments; nothing inside
it changes.**

```{r synthetic-colleague}
set.seed(42)
n <- 400
colleague <- tibble(
  user_id = sprintf("U%04d", seq_len(n)),
  wt      = round(runif(n, 0.5, 2.5), 3),                                   # different weight name
  item_1  = factor(sample(c("Agree", "Neutral", "Disagree"), n, TRUE, c(.5, .3, .2)),
                   levels = c("Agree", "Neutral", "Disagree")),
  item_2  = factor(sample(c("Yes", "No"), n, TRUE), levels = c("Yes", "No")),
  region2 = factor(sample(c("North", "South", "East", "West"), n, TRUE)),
  agegrp  = factor(sample(c("<30", "30-59", "60+"), n, TRUE), levels = c("<30", "30-59", "60+"))
)
colleague_ref <- tibble(
  qvar  = c("item_1", "item_2"),
  qtext = c("Our product is easy to use.", "Would you recommend us?")
)

ct <- weighted_crosstab(
  colleague,
  weight_col      = "wt",
  id_col          = "user_id",
  outcome_pattern = "^item_[0-9]+$",     # different question naming scheme
  reference       = colleague_ref,
  ref_var         = "qvar",
  ref_label       = "qtext"
)

names(ct)                                 # inferred item_1/item_2 × region2/agegrp
ct[["item_1__region2"]]$gt                # wording joined from the renamed reference
```

The same function, unmodified, tabulated a survey with new names and joined wording from a
renamed reference — confirming the reusability requirement.

## Verification

- **Manual validation** (above) recomputes a table cell by two independent paths — base-R
  `weighted.mean()` and `survey::svymean()` — and both match `weighted_crosstab()` exactly.
- **Adversarial testing.** A dedicated agent stress-tested the function: the colleague-rename
  simulation above, plus edge cases (missing/misnamed weight, NA demographics and responses,
  single-level factors, empty subgroups, continuous demographics). It confirmed column
  proportions sum to 100% and surfaced one defect — negative weights silently yielding
  negative percentages — which was fixed by adding a non-negativity check.
